---
title: "Investigating wetland scatter"
author: "Jonathan Bahlmann"
# date: "`r format(Sys.time(), "%d.%m.%Y")`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Data
Set wd and load packages.
```{r, echo=FALSE, warning=FALSE, message=FALSE}
setwd("/home/petra/Praktikum")
library(stars)
library(dplyr)
library(rgdal)
library(gdalUtils)
library(raster)
library(lubridate)
library(sp)
library(ggplot2)
library(zoo)
```

## As stars proxy
Intensity data is handled here.
```{r, echo=FALSE}
# load files as proxy
int_VV <- read_stars("VV_clip.tif", proxy=TRUE)
int_VH <- read_stars("VH_clip.tif", proxy=TRUE)
# assign names
names(int_VV) <- "VV"
names(int_VH) <- "VH"
# transform shape into crs of intensity dataset
shape <- read_sf("antragsfl_16_17_18.shp")
shape <- st_transform(shape, crs=st_crs(int_VV))
```

## Spatial Extent
The desired spatial extent can be set via subsetting: `[extent_object]`.
```{r, echo=FALSE}
# now load extent (polygon) into memory
int_vv <- st_as_stars(int_VV[shape[124,]])
int_vh <- st_as_stars(int_VH[shape[124,]])
```

## Time Dimension
```{r}
dates <- c(ymd("2018-01-03"), ymd("2018-01-15"), ymd("2018-01-27"), ymd("2018-02-08"), ymd("2018-02-20"), ymd("2018-03-04"), ymd("2018-03-16"), ymd("2018-03-28"), ymd("2018-04-09"), ymd("2018-04-21"), ymd("2018-05-03"), ymd("2018-05-15"), ymd("2018-05-27"), ymd("2018-06-08"), ymd("2018-06-20"), ymd("2018-07-02"), ymd("2018-07-14"), ymd("2018-07-26"), ymd("2018-08-07"), ymd("2018-08-19"), ymd("2018-08-31"), ymd("2018-09-12"), ymd("2018-09-24"), ymd("2018-10-06"), ymd("2018-10-18"), ymd("2018-10-30"), ymd("2018-11-11"), ymd("2018-11-23"), ymd("2018-12-05"), ymd("2018-12-17"), ymd("2018-12-29"))
int_vv <- st_set_dimensions(int_vv, 3, values = dates, names = "time")
int_vh <- st_set_dimensions(int_vh, 3, values = dates, names = "time")
```

## Rain
Needs to be freshly warped for some reason, can't be saved (`write_stars`) and then loaded. But: rain_warped.tif is a file with already aggregated rain data (31 time steps instead of 365).
```{r, echo=FALSE}
# rain <- read_stars("rain_warped.tif", proxy=TRUE)
# names(rain) <- "prec"
# rain <- st_warp(rain, int_VV)
# rain <- readRDS("rain_warped_3daysum.rds")
# rain <- rain[shape[124,]]
# rain <- st_set_dimensions(rain, 3, values = dates, names = "time")
```
## Combine
Then combine all three. Give `along = "bands"` so that all attributes are treated as bands.
```{r, echo=FALSE}
# c()
comb <- c(int_vv, int_vh, along="bands")
```
## switch bands to attributes
This is more of a design choice.
```{r}
comb_split <- split(comb, "bands")
names(comb_split) <- c("VV", "VH")
comb_split
```

## Time Series for *all* (a Number of) Pixels
Giving the opportunity of constructing very different results. Depending on whether neighboring pixels or random pixels are selected, similarities and differences are seen.
```{r}
# make first attribute, pixel 10/10 into dataframe template
pix.df <- as.data.frame(comb_split[1,100,100,])[,3:4]
# make data frame for all pixels
for (i in 1:nrow(comb_split)) {
  for (j in 1:ncol(comb_split)) {
    if (is.na(comb_split[1,i,j,1][[1]])) {
    }
    else {
      pix.df <- cbind(pix.df, as.vector(comb_split[1,i,j,][[1]]))
    }
  }
}
# delete template data frame column
pix.df <- pix.df[,c(1,3:ncol(pix.df))]
# assign numbers as names
f <- ncol(pix.df) - 1
names(pix.df) <- c("time", seq(1,f,1))
```

Here we have the possibility to select which pixels are plotted. For example each 37th pixel in the dataframe:

```{r, warning=FALSE}
# initiate plot
p <- ggplot(data=pix.df, aes(x=pix.df$time)) + ylab("VV")
# make sequence of pixel positions to plot
ind <- seq(2,9263, 17)
# loop
for (i in ind) { p <- p + geom_line(aes_string(y = pix.df[,i]), alpha = 0.1) }
p
```