---
title: "Exploring 2019"
author: "Jonathan Bahlmann"
date: "10/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a summary on what I did during my internship at EFTAS in october 2020. The overall topic was to make a connection between SAR backscatter and soil moisture, also including precipitation data. The work can be split into two main approaches: the time series and the classification approach.

# Data

### Packages
```{r libraries, warning=FALSE, message=FALSE}
setwd("/home/petra/Praktikum")
library(stars)
library(dplyr)
library(rgdal)
library(gdalUtils)
library(raster)
library(lubridate)
library(sp)
library(ggplot2)
library(zoo)
library(e1071)
```

### Intensity and Shapefiles
The .tif files each contain data for a year. They are dB - scaled intensity measurements from Sentinel 1 A. All scenes have a offset of 12 days, except for a data gap in January 2017.
```{r loading-int-data}
# load files as proxy
int_VV <- read_stars("./BEWAMO_NEU/S1A-VV-Multilook-Speckle-Gamma0.tif", proxy=TRUE)
# assign names
names(int_VV) <- "VV"
# study area
study_area <- read_sf("study_area.shp")
study_area <- st_transform(study_area, crs=st_crs(int_VV))
# shape
shape <- read_sf("antragsfl_16_17_18.shp")
shape <- st_transform(shape, crs=st_crs(int_VV))

int_VV

```

Dates of scenes are read via `gdalinfo()`.
```{r make-dates}
# make dates
info = gdalinfo("./BEWAMO_NEU/S1A-VV-Multilook-Speckle-Gamma0.tif")
descr = info[grepl(info, pattern = "Description = ")]
descr = gsub(descr, pattern = "Description = ", replacement = "")
dates <- as_date(descr)
```

```{r load-polygon-func}
# function takes a shape as an AOI and a year (2017 or 2018)
loadPolygon <- function(shape, interval) {
  dates <- dates[interval[1]:interval[2]]
  int_vv <- st_as_stars(int_VV[shape], along = "band")
  int_vv <- int_vv[,,,interval[1]:interval[2]]
  int_vv <- st_set_dimensions(int_vv, 3, val = dates, names = "time")
  names(int_vv) <- c("VV")
  return(int_vv)
}
```

```{r}
# explore
area1 <- loadPolygon(study_area[4,], c(25,44))
df <- st_apply(area1, "time", mean, na.rm=TRUE) %>% as.data.frame()
plot(mean ~ time, df, type="l")
```

```{r}
measures <- read.csv("Bodenfeuchte_Rhinluch.csv")
measures <- measures[3:nrow(measures),c(1,6,7,8)]
names(measures) <- c("time", "sensor1", "sensor2", "waterlevel")
measures$time <- dmy(as.character(measures$time))

measures <- aggregate(measures[c("sensor1", "sensor2", "waterlevel")], mean, by = list(Group.date = measures$time))
names(measures) <- c("time", "sensor1", "sensor2", "waterlevel")
measures$sensorMean <- rowMeans(measures[,2:3])

# location is already in WGS84
loc <- st_point(c(12.92506, 52.83639))
loc <- st_sfc(loc, crs=4326)

# extract
point_meas <- st_extract(area1, pts = loc) %>% as.data.frame()

ggplot() +
  geom_line(aes(x = measures$time, y = measures$sensorMean)) +  
  geom_smooth(aes(x = measures$time, y = measures$sensorMean), span = 0.9, color = "darkblue") + 
  geom_line(aes(x = point_meas$time, y = point_meas$VV * 150 + 45)) +
  geom_smooth(aes(x = point_meas$time, y = point_meas$VV * 150 + 45), span = 0.8, color = "red") +
  coord_cartesian(ylim=c(40,60))

ggplot() +
  geom_line(aes(x = measures$time, y = measures$sensorMean)) +  
  geom_smooth(aes(x = measures$time, y = measures$sensorMean), span = 0.25, color = "darkblue") + 
  geom_line(aes(x = point_meas$time, y = point_meas$VV * 150 + 45)) +
  geom_smooth(aes(x = point_meas$time, y = point_meas$VV * 150 + 45), span = 0.3, color = "red") +
  coord_cartesian(ylim=c(40,60))
  
  # scale_color_manual(name = "Förderung", values = c("Sensor mean"="darkblue"))

 #  geom_line(aes(x = measures$time, y = measures$waterlevel / 10 - 50))
```

```{r}
measures <- read.csv("lentzke.csv")
measures <- measures[3:nrow(measures),c(1,6,7,8)]
names(measures) <- c("time", "sensor1", "sensor2", "waterlevel")
measures$time <- mdy(as.character(measures$time))

measures <- aggregate(measures[c("sensor1", "sensor2", "waterlevel")], mean, by = list(Group.date = measures$time))
names(measures) <- c("time", "sensor1", "sensor2", "waterlevel")
measures$sensorMean <- rowMeans(measures[,2:3])

# location is already in WGS84
loc <- st_point(c(12.74090, 52.81763))
loc <- st_sfc(loc, crs=4326)

# extract
point_meas <- st_extract(area1, pts = loc) %>% as.data.frame()

ggplot() +
  geom_line(aes(x = measures$time, y = measures$sensorMean)) +  
  geom_line(aes(x = point_meas$time, y = point_meas$VV * 100 + 43)) +
  coord_cartesian(ylim=c(40,60))
```

```{r}
measures <- read.csv("mf.csv")
measures <- measures[3:nrow(measures),c(1,6,7,8)]
names(measures) <- c("time", "sensor1", "sensor2", "waterlevel")
measures$time <- mdy(as.character(measures$time))

measures <- aggregate(measures[c("sensor1", "sensor2", "waterlevel")], mean, by = list(Group.date = measures$time))
names(measures) <- c("time", "sensor1", "sensor2", "waterlevel")
measures$sensorMean <- rowMeans(measures[,2:3])

# location is already in WGS84
loc <- st_point(c(12.80042, 52.83602))
loc <- st_sfc(loc, crs=4326)

# extract
point_meas <- st_extract(area1, pts = loc) %>% as.data.frame()

ggplot() +
  geom_line(aes(x = measures$time, y = measures$sensorMean)) +  
  geom_line(aes(x = point_meas$time, y = point_meas$VV * 300 + 18))
```