---
title: "Report"
author: "Jonathan Bahlmann"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data

### Packages
```{r libraries, warning=FALSE, message=FALSE}
setwd("/home/petra/Praktikum")
library(stars)
library(dplyr)
library(rgdal)
library(gdalUtils)
library(raster)
library(lubridate)
library(sp)
library(ggplot2)
library(zoo)
library(e1071)
```

### Intensity and Shapefiles
```{r loading-int-data}
# load files as proxy
int_VV <- read_stars("./BEWAMO_NEU/S1A-VV-Multilook-Speckle-Gamma0.tif")
# assign names
names(int_VV) <- "VV"
# water training shapefiles
water_shape_2017 <- read_sf("water (copy).shp")
water_shape_2018 <- read_sf("water.shp")
water_shape_2017_july07 <- read_sf("water_shape_intersec.shp")
# study area
study_area <- read_sf("study_area.shp")
study_area <- st_transform(study_area, crs=st_crs(int_VV))
# shape
shape <- read_sf("antragsfl_16_17_18.shp")
shape <- st_transform(shape, crs=st_crs(int_VV))
# moni
moni <- read_sf("moni_shap.shp")
# double bounce
db <- read_sf("double_bounce.shp")
```

```{r make-dates}
info = gdalinfo("./BEWAMO_NEU/S1A-VV-Multilook-Speckle-Gamma0.tif")
descr = info[grepl(info, pattern = "Description = ")]
descr = gsub(descr, pattern = "Description = ", replacement = "")
dates <- as_date(descr)
```

### Precipitation Data
```{r, warning=FALSE}
# load stars, read and assign dates as dimension
rain_all_2019 <- read_stars("./BEWAMO_NEU/rain_2019_clip.tif")
rainDates2019 <- seq(as.Date("2019-01-01"), as.Date("2019-12-31"), by="days")
rain_all_2019 <- st_set_dimensions(rain_all_2019, 3, values = rainDates2019, names = "time")

# clip to wider bog area, convert to dataframe
study_area_rain <- st_transform(study_area, crs=st_crs(rain_all_2019))
rain_all_2019 <- rain_all_2019[study_area_rain[1,]]
rain.all.2019.df <- as.data.frame(st_apply(rain_all_2019, "time", mean, na.rm = TRUE))

# another method: make vector of all days of a year, assign as dimension
rain_all_2020 <- read_stars("./BEWAMO_NEU/rain_2020_clip.tif")
rainDates2020 <- seq(as.Date("2020-01-01"), as.Date("2020-10-27"), by="days")
rain_all_2020 <- st_set_dimensions(rain_all_2020, 3, values = rainDates2020, names = "time")

rain_all_2020 <- rain_all_2020[study_area_rain[1,]]
rain.all.2020.df <- as.data.frame(st_apply(rain_all_2020, "time", mean, na.rm = TRUE))

# create rain dataframe, rows = days, columns = time, rain measurement
rain_all <- rbind(rain.all.2019.df, rain.all.2020.df)
```

### loadPolygon Function
```{r load-polygon-func}
# function takes a shape as an AOI and a year (2017 or 2020)
loadPolygon <- function(shape, interval) {
  dates <- dates[interval[1]:interval[2]]
  int_vv <- st_as_stars(int_VV[shape], along = "band")
  int_vv <- int_vv[,,,interval[1]:interval[2]]
  int_vv <- st_set_dimensions(int_vv, 3, val = dates, names = "time")
  names(int_vv) <- c("VV")
  return(int_vv)
}
```

# Time Series Approach
The idea of this approach is to extract information about different kinds of scatterers or different soil moisture contents by looking at a time series of values.

### Aggregate Rain
The next step was to plot backscatter and rain data. An important question here is how the rain should be aggregated (since depicting single days might confuse the plot). Different solutions were found. At first, a three day sum was calculated. Then, computations based on what happens between two acquisitions were considered. The following code simply calculates the rain sum between each scene.
```{r aggregate-rain, out.width="100%"}
# cut the bog polygons to our study area
inter_shape <- shape[lengths(st_intersects(shape, study_area[1,])) != 0,]
# cut geometries depending on sponsorship status
spons <- inter_shape[inter_shape$Förderung == "in Förderung",]
nospons <- inter_shape[is.na(inter_shape$Förderung),]
# load polygon, aggregate as mean and convert to data frame
spons.df <- loadPolygon(spons, c(1,44)) %>% st_apply(., "time", mean, na.rm = TRUE) %>% as.data.frame()
nospons.df <- loadPolygon(nospons, c(1,44)) %>% st_apply(., "time", mean, na.rm = TRUE) %>% as.data.frame()
# bind, select, rename
df <- cbind(spons.df, nospons.df) %>% .[,c(1,2,4)]
names(df) <- c("time", "sponsVV", "nosponsVV")

# rain aggregation over both years
agg_rain <- c(rain_all[1,2] + rain_all[2,2] + rain_all[3,2] + rain_all[4,2] + rain_all[5,2] + rain_all[6,2] + rain_all[7,2] + rain_all[8,2]) # to handle the first sum that only has 8 values
for (i in 2:nrow(df)) { # i corresponds to row in S1 acquisitions
  offset <- 8 # first S1 acquisition is 08/01, rain data starts 01/01
  inter <- 12 # 12 days between acquisitions
  start <- (i - 2) * inter + offset + 1 # exclude last from prev sum
  end <- (i - 1) * inter + offset
  sum <- 0 # create sum as 0
  # for the calculated rows of rain data frame
  for (j in start:end) {
    sum <- sum + rain_all[j,2] # add day to day rain
  }
  agg_rain <- c(agg_rain, sum) # append to rain vector
}

df <- cbind(df, agg_rain) # bind to intensity data
names(df) <- c("time", "sponsVV", "nosponsVV", "rain")

ggplot(df, aes(x = time)) + 
  geom_bar(aes(x = time - 6, y = rain), stat = 'identity', fill = "lightblue", alpha = 0.8) + # time - 6 so that rain bars are displayed in between S1 acquisition dates
  geom_line(aes(y = sponsVV * 2000 + 0, color = "in Förderung")) + # scale VV as enhancement of effects
  geom_point(aes(y = sponsVV * 2000 + 0, color = "in Förderung")) +
  geom_line(aes(y = nosponsVV * 2000 + 0, color = "nicht in Förderung")) +
  coord_cartesian(ylim = c(0,300)) + 
  ggtitle("Mean of VV in Bog Polygons and Precipitation") + xlab("Time") +
  ylab("Precipitation in mm/m²") + 
  scale_y_continuous(sec.axis = sec_axis(~. *0.25 -17.5, name = "Intensity in dB")) + 
  scale_color_manual(name = "Förderung", values = c("in Förderung"="blue", "nicht in Förderung"="black")) +
  theme(legend.position = "bottom")
```

Or: build weighted sums of the precipitation between scenes. The following depicts a weight of 1/(# of day before acquisition), so that the day of the acquisition has weight 1, the day before has weight 1/2, 1/3 (...) and so on [code hidden].
```{r rain-weighted, echo=FALSE, out.width="100%"}
# remake df as above, use existing variables
df <- cbind(spons.df, nospons.df) %>% .[,c(1,2,4)]
names(df) <- c("time", "sponsVV", "nosponsVV")

# rain aggregation over both years
agg_rain_weight <- c(rain_all[1,2] + rain_all[2,2] + rain_all[3,2] + rain_all[4,2] + rain_all[5,2] + rain_all[6,2] + rain_all[7,2] + rain_all[8,2]) # to handle the first sum that only has 8 values
for (i in 2:nrow(df)) { # i corresponds to row in S1 acquisitions
  offset <- 10 # first S1 acquisition is 08/01, rain data starts 01/01
  inter <- 12 # 12 days between acquisitions
  start <- (i - 2) * inter + offset + 1 # exclude last from prev sum
  end <- (i - 1) * inter + offset
  sum <- 0 # create sum as 0
  # for the calculated rows of rain data frame
  weight <- 1 # at the start of each 
  for (j in end:start) {
    sum <- sum + (1/weight * rain_all[j,2])
    weight <- weight + 1
  }
  agg_rain_weight <- c(agg_rain_weight, sum) # append to rain vector
}

df <- cbind(df, agg_rain_weight) # bind to intensity data
names(df) <- c("time", "sponsVV", "nosponsVV", "rain")

ggplot(df, aes(x = time)) + 
  geom_bar(aes(x = time - 6, y = rain), stat = 'identity', fill = "lightblue", alpha = 0.8) + 
  geom_line(aes(y = sponsVV * 1000, color = "in Förderung")) +
  geom_point(aes(y = sponsVV * 1000, color = "in Förderung")) +
  geom_line(aes(y = nosponsVV * 1000, color = "nicht in Förderung")) +
  coord_cartesian(ylim = c(0,300)) + 
  ggtitle("Mean of Bog Polygons and Precipitation, Weight: 1 / # of Day") + xlab("Time") +
  ylab("Precipitation in mm/m²") + 
  scale_y_continuous(sec.axis = sec_axis(~. *2/3 - 56/3, name = "Intensity in dB")) + 
  scale_color_manual(name = "Förderung", values = c("in Förderung"="blue", "nicht in Förderung"="black")) +
  theme(legend.position = "bottom")
```
