---
title: "scatter_ts"
author: "Jonathan Bahlmann"
date: "10/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading
Set wd and load packages.
```{r, echo=FALSE, warning=FALSE, message=FALSE}
setwd("/home/petra/Praktikum")
library(stars)
library(dplyr)
library(rgdal)
library(gdalUtils)
library(raster)
library(lubridate)
library(sp)
library(ggplot2)
library(zoo)
```

## Load data
Intensity and rain data is handled here.
```{r, echo=FALSE}
# load files as proxy
int_VV <- read_stars("VV_clip.tif", proxy=TRUE)
int_VH <- read_stars("VH_clip.tif", proxy=TRUE)
# rain must be an aggregated dataset with 31 timesteps
rain <- readRDS("rain_warped_3daysum.rds")
# assign names
names(int_VV) <- "VV"
names(int_VH) <- "VH"
# transform shape into crs of intensity dataset
shape <- read_sf("antragsfl_16_17_18.shp")
shape <- st_transform(shape, crs=st_crs(int_VV))
# monitoring shapes
moni <- read_sf("moni_shap.shp")
water <- read_sf("water.shp")
```

## As stars, Assign date dimension
Set polygonID. Data is clipped accordingly
```{r}
dates <- c(ymd("2018-01-03"), ymd("2018-01-15"), ymd("2018-01-27"), ymd("2018-02-08"), ymd("2018-02-20"), ymd("2018-03-04"), ymd("2018-03-16"), ymd("2018-03-28"), ymd("2018-04-09"), ymd("2018-04-21"), ymd("2018-05-03"), ymd("2018-05-15"), ymd("2018-05-27"), ymd("2018-06-08"), ymd("2018-06-20"), ymd("2018-07-02"), ymd("2018-07-14"), ymd("2018-07-26"), ymd("2018-08-07"), ymd("2018-08-19"), ymd("2018-08-31"), ymd("2018-09-12"), ymd("2018-09-24"), ymd("2018-10-06"), ymd("2018-10-18"), ymd("2018-10-30"), ymd("2018-11-11"), ymd("2018-11-23"), ymd("2018-12-05"), ymd("2018-12-17"), ymd("2018-12-29"))

loadPolygon <- function(shape) {
# This polygon is being looked at: polyID
int_vv <- st_as_stars(int_VV[shape])
int_vh <- st_as_stars(int_VH[shape])
int_vv <- st_set_dimensions(int_vv, 3, val = dates, names = "time")
int_vh <- st_set_dimensions(int_vh, 3, val = dates, names = "time")
# c()
comb <- c(int_vv, int_vh, along="bands")
# switch bands to attributes, this is more of a design choice.
comb_split <- split(comb, "bands")
names(comb_split) <- c("VV", "VH")
return(comb_split)
}

comb_split10 <- loadPolygon(water[10,])
```

## plot rain of complete AOA
```{r, echo=FALSE, eval=FALSE}
rain_all <- read_stars("rain_clip.tif")
all_days <- seq(as.Date("2018-01-01"), as.Date("2018-12-31"), by="days")
study_area_rain <- st_transform(study_area, crs=st_crs(rain_all))
# rain_all <- rain_all[shape_rain]
rain_all <- st_set_dimensions(rain_all, 3, values = all_days, names = "time")
# cut
rain_all <- rain_all[study_area_rain[3,]]
rain.all.df <- as.data.frame(st_apply(rain_all, "time", mean, na.rm = TRUE))
df <- rain.all.df[1:30,]
ggplot(df, aes(x = time, y = mean)) +
  geom_bar(stat='identity')# +
 #  geom_vline(xintercept = as.numeric(as.Date("2018-02-20")), linetype=3) +
#   geom_vline(xintercept = as.numeric(as.Date("2018-03-04")), linetype=3) +
#   geom_vline(xintercept = as.numeric(as.Date("2018-03-16")), linetype=3)


```

## Plotting
```{r}
# whole year, both pols
# VV
plot(comb_split10[1,,,])
# VH
plot(comb_split10[2,,,])
```

Function to plot a normalized change ratio.

```{r, echo=FALSE, warning=FALSE}
plotCD <- function(starsObj, Attr, start, end) {
  end <- end - 1
  colo <- viridisLite::inferno(60)
  brk <- seq(-0.3,0.3,0.01)
  for (i in start:end) {
    cd <- (starsObj[Attr,,,i] - starsObj[Attr,,,i+1]) / (starsObj[Attr,,,i] + starsObj[Attr,,,i+1])
    title <- paste0("Norm. Change Ratio ", dates[i], " to ", dates[i+1])
    image(cd, col = colo, breaks = brk, main = title) # maxpixels=ncell(cd), main=title)
  }
}
```

```{r, fig.show="hold", out.width="25%", warning=FALSE}
# VV, from 1 to 31
plotCD(comb_split10, 1, 1, 5)
```

Function to plot a timeseries of pixel for a certain polygon
```{r, echo=FALSE, warning=FALSE}
ploTShape <- function(starsObj, polygon, polygon2 = NULL) {
# initiate
pix.df.tmp <- as.data.frame(comb_split10[1,18,18,])[,3:4]
pix.df <- pix.df.tmp
pix.df.2 <- pix.df.tmp
# clip
ext <- starsObj[polygon]
# make data frame for all pixels
for (i in 1:nrow(ext)) {
  for (j in 1:ncol(ext)) {
    if (is.na(ext[1,i,j,1][[1]])) {
    }
    else {
      pix.df <- cbind(pix.df, as.vector(ext[1,i,j,][[1]]))
    }
  }
}

if(!is.null(polygon2)) {
  ext2 <- starsObj[polygon2]
  # make data frame for all pixels
  for (i in 1:nrow(ext2)) {
    for (j in 1:ncol(ext2)) {
      if (is.na(ext2[1,i,j,1][[1]])) {
      }
      else {
        pix.df.2 <- cbind(pix.df.2, as.vector(ext2[1,i,j,][[1]])) 
      }
    }
  }
  pix.df.2 <- pix.df.2[,c(1,3:ncol(pix.df.2))]
  f <- ncol(pix.df.2) - 1
  names(pix.df.2) <- c("time", seq(1,f,1))
}
else {}

# delete template data frame column
# pix.df <- pix.df[,c(1,3:ncol(pix.df))]
# assign numbers as names
f <- ncol(pix.df) - 2
names(pix.df) <- c("time", "prec", seq(1,f,1))

  # initiate plot
  p <- ggplot(data=pix.df, aes(x=pix.df$time))
    # ylab("Precipitation") + 
    # geom_bar(aes(y=pix.df$prec), stat='identity') +
    # scale_y_continuous(sec.axis = sec_axis(~. + 22, name = "Intensity + 22"))
  
  # make sequence of pixel positions to plot
  ind <- seq(3,ncol(pix.df), 7)
  # loop
  for (i in ind) { p <- p + geom_line(aes_string(y = pix.df[,i]), alpha = 0.2) }
  if(!is.null(polygon2)) {
      inde <- seq(2, ncol(pix.df.2), 7)
      for (j in inde) { p <- p + geom_line(aes_string(y = pix.df.2[,j]), alpha = 0.1, color="red") }
  } else{}

  p
}
```

#### Compare Two Suspected Water Patches
```{r, warning=FALSE}
# both are thesis: water
ploTShape(comb_split10, moni[8,], moni[11,])
```

#### Water (black) and "Random" Field (red)
```{r, warning=FALSE}
ploTShape(comb_split10, moni[10,], moni[9,])
```

#### Water (black) and other field (red)
```{r, warning=FALSE}
# ploTShape(comb_split10, moni[4,], moni[6,])
```

#### polygon - water shape and water shape itself
```{r, warning=FALSE}
plotbox <- function(starsObj, date, polygon1, polygon2) {
  poly1 <- starsObj[polygon1]
  poly2 <- starsObj[polygon2]
  poly1.df <- as.data.frame(poly1[,,,date])
  poly1.df$cat <- "polygon 1"
  poly2.df <- as.data.frame(poly2[,,,date])
  poly2.df$cat <- "polygon 2"
  poly <- rbind(poly1.df, poly2.df)
  ggplot(poly, aes(x=poly$cat, y=poly$VV)) + geom_boxplot()
}

plotbox(comb_split10, 1, moni[2,], shape[124,])
# plotbox(comb_split10, 3, moni[2,], shape[124,])
# plotbox(comb_split10, 6, moni[2,], shape[124,])
```

```{r, fig.show="hold", out.width="33%", warning=FALSE}
plotthree <- function(obj, start, end) {
colo = viridisLite::inferno(85)
breaks = seq(-25, -8, 0.2)
threshold = -15

for (i in start:end) {
  plot(obj[1,,,i])
  image(obj[1,,,i], col = colo, breaks = breaks)
  image(obj[1,,,i], col = c("white", "black"), breaks = c(-22, threshold, 0))
}
}

# plotthree(comb_split10, 1, 7)
```

## Look at step 6 of other polygons
```{r, fig.show="hold", out.width="50%", warning=FALSE}
comb_split124 <- loadPolygon(moni[8,])
# VV
plotCD(comb_split10, 1, 9, 11)
```
```{r, fig.show="hold", out.width="50%", warning=FALSE}
# VH
plotCD(comb_split10, 2, 9, 11)
```
```{r, fig.show="hold", out.width="33%", warning=FALSE}
# plots: stars-plot, stars-image with unified scale, masked image
plotthree(comb_split10, 1, 31)
```

```{r, fig.show="hold", out.width="50%", warning=FALSE}
comb_split123 <- loadPolygon(moni[9,])
# VV
plotCD(comb_split123, 1, 9, 11)
```
```{r, fig.show="hold", out.width="50%", warning=FALSE}
# VH
plotCD(comb_split123, 2, 9, 11)
```
```{r, fig.show="hold", out.width="33%", warning=FALSE}
# plots: stars-plot, stars-image with unified scale, masked image
plotthree(comb_split123, 9, 11)
```

```{r, fig.show="hold", out.width="50%", warning=FALSE}
comb_split112 <- loadPolygon(moni[10,])
# VV
plotCD(comb_split112, 1, 9, 11)
```
```{r, fig.show="hold", out.width="50%", warning=FALSE}
# VH
plotCD(comb_split112, 2, 9, 11)
```
```{r, fig.show="hold", out.width="33%", warning=FALSE}
# plots: stars-plot, stars-image with unified scale, masked image
plotthree(comb_split112, 9, 11)
```

```{r, fig.show="hold", out.width="50%", warning=FALSE}
comb_split108 <- loadPolygon(moni[11,])
# VV
plotCD(comb_split108, 1, 9, 11)
```
```{r, fig.show="hold", out.width="50%", warning=FALSE}
# VH
plotCD(comb_split108, 2, 9, 11)
```
```{r, fig.show="hold", out.width="33%", warning=FALSE}
# plots: stars-plot, stars-image with unified scale, masked image
plotthree(comb_split108, 9, 11)
```

