---
title: "Characterize by Change"
author: "Jonathan Bahlmann"
date: "10/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, echo=FALSE, warning=FALSE, message=FALSE}
setwd("/home/petra/Praktikum")
library(stars)
library(dplyr)
library(rgdal)
library(gdalUtils)
library(raster)
library(lubridate)
library(sp)
library(ggplot2)
library(zoo)
library(e1071)
```

```{r loading-data, echo=FALSE}
# load files as proxy
int_VV_2017 <- read_stars("VV_2017_clip.tif", proxy=TRUE)
int_VH_2017 <- read_stars("VH_2017_clip.tif", proxy=TRUE)
int_VV_2018 <- read_stars("VV_clip.tif", proxy=TRUE)
int_VH_2018 <- read_stars("VH_clip.tif", proxy=TRUE)
# assign names
names(int_VV_2017) <- "VV"
names(int_VH_2017) <- "VH"
names(int_VV_2018) <- "VV"
names(int_VH_2018) <- "VH"
# monitoring waters
water_shape_2017 <- read_sf("water (copy).shp")
water_shape_2018 <- read_sf("water.shp")
water_shape_2017_july07 <- read_sf("water_shape_intersec.shp")
# study area
study_area <- read_sf("study_area.shp")
# shape
shape <- read_sf("antragsfl_16_17_18.shp")
shape <- st_transform(shape, crs=st_crs(int_VV_2018))
# moni
moni <- read_sf("moni_shap.shp")
# double bounce
db <- read_sf("double_bounce.shp")

# water_general <- water_shape_2017[lengths(st_intersects(water_shape_2017, water_shape_2018)) != 0,]
```

## Characteristics

Bands are:

* change count 
  + how often does the pixel change? -> count of change
  + ranges from 0 to # of time steps
  + higher values with more often change
* change frequency 
  + how stable is the pixel? -> mean of no-change-streaks
  + ranges from 1 to # of time steps
  + higher values correspond to longer periods of no change
  + switch to the maximum of these periods?
* change range 
  + what magnitude does the change have? -> difference of max and min
  + range dependent on max difference
  + higher values indicate *one* high contrast event
* mean
  + range calculated
  + higher values for a higher mean backscatter
* threshold count
  + count over the time period of one specific range of values -> how often were the pixels between a and b
  + max is # of time steps
  + high values depict high counts
  
```{r, warning=FALSE, fig.show="hold", out.width="50%"}
sta <- read_stars("charac_five_attr/all_2017-01-08_2018-12-29.tif")
study_area <- st_transform(study_area, crs=st_crs(sta))
sta_small <- sta[study_area[1,]]
plot(sta_small[,,,1], main = "Change Count")
plot(sta_small[,,,2], main = "Change Frequency")
plot(sta_small[,,,3], main = "Change Range")
plot(sta_small[,,,4], main = "Mean")
plot(sta_small[,,,5], main = "Threshold Count")
```

## Use Case Example: Count and Range

* pixels that change often but not by much: high count and low range -> high +
* pixels that change often and mostly with high range: high count, high range -> ~ 0
* pixels that change not so often but if they do, they do by a lot: low count, high range -> low -
* pixels that change not so often and if they do, they only change slightly: low count, low range -> ~ 0

```{r, warning=FALSE}
# conversion to raster scales to 0:255
sta_small_count <- as(sta_small[,,,1], "Raster")
sta_small_range <- as(sta_small[,,,3], "Raster")
# NDI
ras <- (sta_small_count - sta_small_range) / (sta_small_count + sta_small_range)
# color
colo = viridisLite::inferno(20)
plot(ras, col = colo)
```

## Full Extent RGB: R = Count, G = Frequency, B = Range

red -> much change, green -> much stability over time, blue -> high change amplitues
```{r, warning=FALSE, out.width="100%"}
ras <- raster::stack("charac_five_attr/all_2017-01-08_2018-12-29.tif")
plotRGB(ras, r=1, g=2, b=3, stretch = "lin")
```

## Full Extent RGB: R = Count, G = Frequency, B = Threshold Count

red -> much change, green -> much stability over time, blue -> high water threshold counts
```{r, warning=FALSE, out.width="100%"}
ras <- raster::stack("charac_five_attr/all_2017-01-08_2018-12-29.tif")
ras <- crop(ras, study_area[1,])
ras <- mask(ras, study_area[1,])
plotRGB(ras, r=1, g=2, b=5, stretch = "lin")
```

## Further Possibillities

* downscale / filter the tiles or the input rasters to account for speckle
* band naming